---
title: "HW2"
author: "Yuyi Yang"
date: "`r Sys.Date()`"
output: html_document
---

# 1. Import class one survey data from our Github site (https://github.com/kijohnson/ADA) (The dataset is called ‘Class 1 Survey Fall 2022.csv’), calling the R dataframe that you create C1survey.
```{r}
pacman::p_load(dplyr, readr)
C1survey<-read_csv("https://raw.githubusercontent.com/kijohnson/ADA/master/Class%201%20Survey%20Fall%202022.csv")
```
# 2. Determine: a. the number of observations (i.e. the number of people who filled out the survey) and b. the number of variables in the dataframe.
```{r}
dim(C1survey)
# 42 obs. of 27 variables
```
# 3. a. Rename the column variables to something shorter and that is descriptive of what the variable is about (for example like_dogs for the ‘Do you like dogs?’ question variable) and b. Display that they are renamed.
```{r}
names(C1survey)[1:27]<-c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three","public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city", "highest_educ_level")
colnames(C1survey)
```
# 4. Write code to determine and display the number of factor, integer, numerical, and character variables there are in the C1survey dataset.
```{r}
table(sapply(C1survey, class))
# 23 character, 4 numeric variabls
```
# 5. a. Using code, check the bday and bmonth variables for any unusual or missing values. If you find any, b. describe what you will do with the unusual values in a comment before or immediately following the code that does it, and c. after you have cleaned up any unusual values, find the median bday and month. *Note: please make sure that your bday and month variables are numeric before doing statistics
```{r}
# check the variables
C1survey$bday
C1survey$bmonth

# bday recoding
class(C1survey$bday) 
#check that it is coded as character class
C1survey$bday<-ifelse(C1survey$bday=="13-Aug-97", "13", C1survey$bday)
C1survey$bday<-ifelse(C1survey$bday=="June 8th", "8", C1survey$bday)
#convert to numeric for statistics
C1survey$bday<-as.numeric(C1survey$bday)
#check values 
C1survey$bday

#bmonth recoding
class(C1survey$bmonth)
C1survey$bmonth[C1survey$bmonth=="June 8th, 1998"]<-"6"
C1survey$bmonth[C1survey$bmonth=="October"]<-"10"
C1survey$bmonth[C1survey$bmonth=="August"]<-"8"
C1survey$bmonth[C1survey$bmonth=="March"]<-"3"
C1survey$bmonth<-as.numeric(C1survey$bmonth)
C1survey$bmonth #check values 

median(C1survey$bday, na.rm=TRUE) #15
median(C1survey$bmonth, na.rm=TRUE) #7
# or
summary(C1survey$bday)
summary(C1survey$bmonth)
```
# 6. a. Create a new variable called bseason that gives the season according to Northern Meteorological season in which respondents were born (winter=12/1 to 2/29, spring=3/1 to 5/31, summer=6/1 to 8/31, fall=9/1 to 11/30). b. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct c. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code.
```{r}
C1survey<-C1survey %>%
  mutate(bseason=case_when(bmonth %in% c(12,1,2) ~"Winter",
                           bmonth %in% c(3,4,5) ~"Spring",
                           bmonth %in% c(6,7,8) ~"Summer",
                           bmonth %in% c(9,10,11) ~"Fall"))
C1survey$bseason
C1survey$bseason_f<-factor(C1survey$bseason, labels=c("Fall", "Spring", "Summer", "Winter"))
table(C1survey$bmonth, C1survey$bseason)

tablesum<-addmargins(table(C1survey$bmonth, C1survey$bseason), 1)
tablesum

#Births by season: Winter=9, Spring=8, Summer=13, Fall=11
```
# 7. Pick your favorite variable to analyze, come up with a question you want to answer with that variable, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #’s).
```{r}
# Question: How many classmates like the March most in our class?

# recode those who like March most as 1, others as 0.
C1survey$like_March <- ifelse(C1survey$fav_month == "March",1,0)
class(C1survey$like_March)

# Use the factor function to change from a numerical variable to a factor variable and label levels
C1survey$like_March <- factor(C1survey$like_March, levels=c(0,1), 
              labels = c("Like other months most", "Like March most")) 

# make sure it is a factor
class(C1survey$like_March)

# get the number of each level
table(C1survey$like_March)
#There are 3 classmates like March most in the class.
```

